Racial and Ethnic Differences in ETI: Outcomes

### 2024-07-07 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
need desktop for some of this I think

```{r setup}
#| include: false
# Libraries
library(tidyverse)
library(lmerTest)
library(performance)
library(segmented)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

1. Lung function percent predicted values < 20% or > 150% were excluded.
2. Limited to 3 years before and 2 years after ETI start.
  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
3. Hospitalizations and home IV episodes excluded.

@@

<!-- # Segmented models -->

<!-- The following models are adjusted for age group and insurance status. First, a Gaussian mixed model is fit with random intercept for patient ID. Then, the `segmented.lme` function is applied to this mixed model, with no pre-specified starting values and `npsi = 2`. The change points are treated as random effects and allowed to vary within a person. -->

<!-- ## FEV1 -->

<!-- Based on visual inspection of the LOESS plots, I expect there to be a change in slope at around 0 and maybe a second one at around 1 year. -->

<!-- ```{r} -->
<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_single_cp_fit.RData") -->
<!-- plot(fev1_single_cp_fit) -->
<!-- summary(fev1_single_cp_fit) -->
<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/segmented/fev1_double_cp_fit.RData") -->
<!-- plot(fev1_double_cp_fit) -->
<!-- summary(fev1_double_cp_fit) -->
<!-- ``` -->

<!-- ## A Bayesian approach -->

<!-- ### FEV1 -->

<!-- ```{r} -->
<!-- #| eval: false -->
<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_single_cp_fit.RData") -->
<!-- load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/fev1_double_cp_fit.RData") -->
<!-- ``` -->

<!-- - Is the second change point due to who is being collected after one year out? -->

<!-- Look at 3 years before and two after -->

<!-- 2 years for 12 and up -->
<!-- 1.5 years for 6 -->

>>>>

```{r setup}
#| include: false
# Libraries
library(tidyverse)
library(lmerTest)
library(splines)
library(ggeffects)
library(performance)
library(segmented)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

1. Lung function percent predicted values < 20% or > 150% were excluded.
2. Limited to 3 years before and 2 years after ETI start.
  - Based on `Modulator_trikafta_first_date` variable provided by CFFPR.
3. Hospitalizations and home IV episodes excluded.
4. Years of pregnancy excluded (we only have annualized pregancy data, so don't know exact dates for this).

@@

# Segmented models

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.

```{r}
invisible(lapply(continuous_outcomes), function(o) {
  f <- as.formula(paste0(
    o,
    "~ age_group + Insurance + bs(Days, knots = 0, degree = 1) * Race + (1 | eDWID)"
  ))
  m = lmer(f,data = encounter)
  predict_response(model = m, terms = c("Days[all]", "Race")) %>% plot()
})

```
